---
title: "Aides Drug"
author: "Daryl Perry"
date: "Fall 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r}
DrugUsers<- filter(triUmp,DrugUser=="yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Thru the use of data files the effective use of a preventive aids drug undergoing clinical trials will be observed from the results of graphic,numerical and inferential data results which will be extracted for greater clarity of the data analysis.The Selected focus filter will be on the triUmp, subdivision labeled DrugUser. 

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods
The method of choice is graphical and numerical tools to aid in providing evidence needed to make an
 analysis of the data to support or reject the null hypotheses.The barchartGc and the Chi test in addition to the Fisher test. 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results
```{r eval=FALSE}
DrugUsers <- DrugUsers %>%
mutate(across(where(is.character), as.factor))
```
<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results
```{r}
barchartGC(~infected + treatment,data=DrugUsers)
barchartGC(~infected + treatment,data=DrugUsers, type="percent")

barchartGC(~treatment + infected,data=DrugUsers)
barchartGC(~treatment + infected,data=DrugUsers, type="percent")


```
The barcharts above shows information that requires careful coding.If the order of the query is reversed the accurate reading of the information in restricted due to clarity and size of the graph data. observe the charts above and pay particular attention to the size and coding request order.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~treatment + infected,data=DrugUsers)
rowPerc(table1)
colPerc(table1)
```
This data shows us that the graphic data results given by the barchartGc



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
This text (Fisher test shows the odds ratio being 1.5439 which indicates the times you are likely to get aids 
while taking the drug)


# Discussion and Conclusion




<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
